---
title: "PA1_template"
author: "Ngan"
date: "Sunday, December 14, 2014"
output: html_document
---

Loading and preprocessing the data

1. load the data
#head(activity)
#names(activity)

```{r}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
```

2. preprocess the data

```{r, eval=T, echo=TRUE}
library(lattice)
activity$date <- as.Date(activity$date, "%Y-%m-%d")

```

What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r, eval=T}
StepsTotal <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
hist(StepsTotal$steps, main = "Total steps by day", xlab = "day", col = "red")

```

2. Calculate and report the mean and median total number of steps taken per day
```{r, eval=T}
mean(StepsTotal$steps)
median(StepsTotal$steps)
```
What is the average daily activity pattern?
```{r, eval=T}

steps <- rep(NA, 61)
day <- rep("NA", 61)
stepsday <- tapply(activity$steps, activity$date, sum, na.rm = T)
length(stepsday)
for (i in 1:61) {
  steps[i] <- stepsday[[i]]
  day[i] <- names(stepsday)[i]
}
df <- data.frame(day, steps)
head(df)
hist(df$steps, main = "Total steps by day", xlab = "day", col = "green")
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, eval=T}

plot(row.names(time_series), time_series, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Average number of steps taken", 
     col = "red")

max_interval <- which.max(time_series)
names(max_interval)

```

Imputing missing values
```{r}

StepsAverage <- aggregate(steps ~ interval, data = activity, FUN = mean)

fillNA <- numeric()
for (i in 1:nrow(activity)) {
    obs <- activity[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}

activity_NA <- sum(is.na(activity))
activity_NA

new_activity <- activity
new_activity$steps <- fillNA

StepsTotal2 <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)

hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")

mean(StepsTotal2$steps)

median(StepsTotal2$steps)

```
Are there differences in activity patterns between weekdays and weekends?

```{r}
day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
  if (day[i] == "Saturday") {
    daylevel[i] <- "Weekend"
  } else if (day[i] == "Sunday") {
    daylevel[i] <- "Weekend"
  } else {
    daylevel[i] <- "Weekday"
  }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")

xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")

```

